Applying Radial Basis Networks and Markov Chains for on-line detection of concept drift in non-stationary environments

### 2019-10-27 tex/article/article.tex
Abstract done

\begin{abstract}
Most real-world problems experience a phenomenon known as concept drift,
a change in data distribution that can affect the system performance.
%
However, the majority of drift detection methods are unsuited for non-stationary environments with data streams.
These algorithms or require the correct labeling of data - infeasible in these settings,
or do not match the severe response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property
and uses Markov Chains to model the drifts transitions.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A problem with impact in different areas of knowledge,
since many behavioral experiments use eye-tracking information as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
\end{abstract}

@@

The rest of the paper is organized as follows:
Section 2 describes the concept drift phenomenon and the main detection techniques;
Section 3 presents the fixations and saccades detection problem;
Section 4 describes the RBFChain algorithm and its pseudo-code;
Section 5 shows the experiment configuration for synthetic datasets  and results obtained;
Section 6 presents the examination setup with a real-world eye-tracking problem and observed outcomes; and, finally, Section 7 provides conclusions
and discusses future work.

>>>>

\begin{abstract}
The amount of information produced by computer systems has grown sharply in recent decades.
Most of this information is presented in the form of uninterrupted and potentially infinite sequences known as data streams.
These streams are usually produced in non-stationary environments,
in which the data distribution can change over time,
possibly deteriorating the system performance.
In the literature,
this phenomenon is named concept drift.
%
Nevertheless,
most drift detection methods are unsuited for non-stationary environments with data streams.
These algorithms or require the correct labeling of data - infeasible in these settings,
or do not match the strict response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property
and uses Markov Chains to model the drifts transitions.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A problem with impact in different areas of knowledge,
since many behavioral experiments use eye-tracking information as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time,
with high accuracy, noise tolerance, and using limited resources.
\end{abstract}

@@

The rest of the paper is organized as follows:
Section 2 describes the concept drift phenomenon and the main detection techniques;
Section 3 presents the eye-tracking problem, with a focus on the identification of fixations and saccades;
Section 4 describes the RBFChain algorithm and its pseudo-code;
Section 5 shows the experiment configuration for synthetic datasets and the obtained results;
Section 6 presents the examination setup with a real-world eye-tracking problem and observed outcomes; and, finally, Section 7 provides conclusions
and discusses future work.
